\chapter{Introduction}
\label{chap:introduction}
This chapter offers a brief introduction to this diploma thesis. In addition, it gives an
overview of this report at the end of the chapter. 

\section{Motivation}
\label{sec:motivation}
E-Books gained a great deal of popularity over the past few years. ...(E-Books today)...  
E-Books are especially useful in a university environment, where popular textbooks 
are not always available due to a small number of printed copies. 
However, due to restrictions in the German copyright law 
(Urheberrechtsgesetz, UrhG)~\cite{wurhg} university libraries are currently unable
to distribute E-Books to students the same way they distribute printed books. The reason for that is the 
lack of standard application and tools for distributing E-Books, that meet all of requirements of 
UrhG for public libraries. 
Indeed, most of the popular E-Book readers today such as ...(Adobe, etc)...  
do not meet many of requirements of UrhG for public libraries,
the most significant of which is the restriction to copy and read the E-Book outside the
 library building. It should be noted that, there are some exceptions to 
 this rule, such as the ability for a student to print up to 75 percent of a book over time.
 
The main focus of this thesis is to implement a viewer that meets the requirements of the UrhG. 
As a basis for our project we use a tool called \emph{ELBS Viewer}, which was
developed as a part of a previous project at the W\"urzburg University, 
Department of Computer Science, 
Char for Computer Science II. 
The purpose of this tool is to provide a way for students to view books online, which were  
digitalized form the standard collection of the University Library Würzburg. 
In addition, the viewer prohibited the copying of a book or parts of a book.
\section{Project Goals}
\label{sec:goals}
As a part of this thesis the ELBS Viewer should be extended in such way that it can meet
the criteria of the German copyright law 
(Urheberrechtsgesetz, UrhG), in particular Article 52b and Article 53~\cite{wurhg}:

\begin{description}
\item[UrhG, § 52b:] Wiedergabe von Werken an elektronischen Leseplätzen in öffentlichen Bibliotheken, 
Museen und Archiven

Zulässig ist, veröffentlichte Werke aus dem Bestand öffentlich zugänglicher 
Bibliotheken, Museen oder Archive, die keinen unmittelbar oder mittelbar wirtschaftlichen oder 
Erwerbszweck verfolgen, ausschließlich in den Räumen der jeweiligen Einrichtung an eigens dafür 
eingerichteten elektronischen Leseplätzen zur Forschung und für private Studien zugänglich zu machen,
soweit dem keine vertraglichen Regelungen entgegenstehen. Es dürfen grundsätzlich nicht mehr Exemplare 
eines Werkes an den eingerichteten elektronischen Leseplätzen gleichzeitig zugänglich gemacht werden, 
als der Bestand der Einrichtung umfasst. Für die Zugänglichmachung ist eine angemessene Vergütung zu 
zahlen. Der Anspruch kann nur durch eine Verwertungsgesellschaft geltend gemacht werden. 


\item[UrhG, § 53:] Vervielfältigungen zum privaten und sonstigen eigenen Gebrauch

(1) Zulässig sind einzelne Vervielfältigungen eines Werkes durch eine 
natürliche Person zum privaten Gebrauch auf beliebigen Trägern, sofern 
sie weder unmittelbar noch mittelbar Erwerbszwecken dienen ... . 
Der zur 
Vervielfältigung Befugte darf die Vervielfältigungsstücke auch durch einen 
anderen herstellen lassen, sofern dies unentgeltlich geschieht oder es sich 
um Vervielfältigungen auf Papier oder einem ähnlichen Träger mittels beliebiger 
photomechanischer Verfahren oder anderer Verfahren mit ähnlicher Wirkung handelt. 

(2) Zulässig ist, einzelne Vervielfältigungsstücke eines Werkes herzustellen oder herstellen zu lassen

1. zum eigenen wissenschaftlichen Gebrauch, wenn und soweit die Vervielfältigung zu 
diesem Zweck geboten ist und sie keinen gewerblichen Zwecken dient,

2. zur Aufnahme in ein eigenes Archiv, wenn und soweit die Vervielfältigung zu 
diesem Zweck geboten ist und als Vorlage für die Vervielfältigung ein eigenes Werkstück benutzt wird,

...

4. zum sonstigen eigenen Gebrauch,
\begin{description}
    \item a) wenn es sich um kleine Teile eines erschienenen Werkes oder um einzelne Beiträge handelt, die in 
    Zeitungen oder Zeitschriften erschienen sind,

    \item b) wenn es sich um ein seit mindestens zwei Jahren vergriffenes Werk handelt. 
\end{description}
(4) Die Vervielfältigung ...

b) eines Buches oder einer Zeitschrift, wenn es 
sich um eine im wesentlichen vollständige Vervielfältigung handelt, ist ... zulässig ... 
unter den Voraussetzungen des Absatzes 2 Satz 1 Nr. 2 oder zum eigenen Gebrauch, wenn es 
sich um ein seit mindestens zwei Jahren vergriffenes Werk handelt. 

(6) Die Vervielfältigungsstücke dürfen [vom Endnutzer] weder 
verbreitet noch zu öffentlichen Wiedergaben benutzt werden.

... 
\end{description}


Another major part of this thesis is the development of a separate Web-based system, 
which can enable the
University Library Würzburg to easily redistribute digitalized documents of any kind
to other parties for research and/or commercial purposes. 
Furthermore, the system should be able to manage user data and documents in such way that 
documents can be made available only to certain user groups and for limited period of
time or number of downloads. In addition, the system 
should be integrated with the already existing workflow-system \emph{WUEBDIFON},
which administrates the process of digitalizing documents.  

\section{Requirements}
In the following we present some of the most important requirements for the software 
upon which this thesis is based.

Due to the fact that our implementation has to be meet the requirements of the UrhG....
-- 

\section{Brief overview of the ELBS Viewer}
In order
for the reader to better grasp the concept of the ELBS viewer, we give a brief over of its
functionality in this section. 

In the remaining portion of the report we distinguish between the two versions of the ELBS E-Book viewer, thus 
we refer to the
original version as ELBS 1.0; and to the improved version discussed in this report as
ELBS 1.1. 


-- 

\section{Comparison with Other Systems}

-- ELBS vs. ??? Adobe state of the art
-- Download Portal vs DoXchanger 

\section{Organization of this Report}
\label{sec:organization}



