\chapter{Approach}
\label{chap:approach}

Approach...

\section{Design concepts} 
\label{sec:approach-design}
\subsection{Platform choice}

- Comparing different environments:

- Web-based vs. standalone application. 

- Java applets vs. Flash vs. AJAX

\subsection{AJAX}
\label{sec:ajax}
- AJAX Philosophy

- What is an asynchronous communication with the server (+ diagram).

- Simple examples with AJAX 

- Problems with AJAX 

- Different browser often requires different code.

- Code management problems.

- Not very good tools for developing with JavaScript.

\subsection{GWT} 
\label{sec:gwt}
- GWT Philosophy

- What is GWT (Google Web Toolkit)?

- Why is GWT better than pure JavaScript.

- Debugger.

- Simple exmles with GWT

\subsection{Limitations of JavaScript and GWT}

- Single thread environment.

- Why do we need a web server

- No database nor file system connectivity.

- Cross site scripting.

\subsection{Server communication}

- How is the communication with the server done 
-> asynchronous, with XML as data exchange format

- Choice for programming language for server-side scripting -> PHP - why?

- Choice for database system -> MySQL - why?

- Outlining that the client side is server independent and other 
implementations of the server side scripting and database schemas are possible, 
without the need of changing the client program.

\subsection{Other design issues}
- Assignments 
 
- Initial approach -> Generating assignments and other suggestions which were 
not implemented. Why?

- Final design ->Environment where user can create assignments and solve them 
and have a way to communicate with each other.
